---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story "Lawmakers expensed millions in 2023 under new program that doesn’t require receipts" (https://www.washingtonpost.com/politics/2024/06/04/house-representatives-expenses-receipts-lodging/)

## Load libraries

Loading required libraries for this analysis.

```{r}
options(scipen = 999)
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
jan_mar <- read.csv("Downloads/reverse_engineering/data/JAN-MAR-2023-SOD-DETAIL-GRID-FINAL.csv")
apr_jun <- read.csv("Downloads/reverse_engineering/data/APRIL-JUNE-2023-SOD-DETAIL-GRID-FINAL.csv")
jul_sep <- read.csv("Downloads/reverse_engineering/data/JULY-SEPTEMBER-2023-SOD-DETAIL-GRID-FINAL.csv")
oct_dec <- read.csv("Downloads/reverse_engineering/data/OCT-DEC-2023-SOD-DETAIL-GRID-FINAL.csv")
```

```{r}
apr_jun$BUDGET.OBJECT.CODE <- as.character(apr_jun$BUDGET.OBJECT.CODE)
oct_dec$BUDGET.OBJECT.CODE <- as.character(oct_dec$BUDGET.OBJECT.CODE)
jul_sep$BUDGET.OBJECT.CODE <- as.character(jul_sep$BUDGET.OBJECT.CODE)
jan_mar$BUDGET.OBJECT.CODE <- as.character(jan_mar$BUDGET.OBJECT.CODE)
```

```{r}
FY2023 <- bind_rows(list(apr_jun, jan_mar, jul_sep, oct_dec))
```

```{r}
rep_expenses <- FY2023 |>
  mutate(
    PERFORM.START.DT = as.Date(PERFORM.START.DT, format = "%d-%b-%y"),
    PERFORM.END.DT = as.Date(PERFORM.END.DT, format = "%d-%b-%y"),
    TRANSACTION.DATE = as.Date(TRANSACTION.DATE, format = "%d-%b-%y")
  )
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Rep. Jack Bergman (R-Mich.) was the program’s top spender last year. He was reimbursed more than $32,000 for lodging and nearly $12,000 for meals in 2023, according to data released by the House as of Tuesday.
* **Analysis summary**: For Rep. Jack Bergman we found that he spent $49,695.49 in lodging in 2023, which is well over the $32,000 he was quoted for in the article. However, we were only able to reach a maximum of $4,388 for meals, and only if we added in expenses from the food and beverage category. We suspect that some other expenses were reclassified to meals for the total of $12,000

```{r}
jack_bergman <- rep_expenses |>
  filter(ORGANIZATION == "2023 HON. JACK BERGMAN")

jack_bergman |>
  mutate(MONTH = month(as.Date(PERFORM.END.DT, format = "%d-%b-%y"))) |>
  group_by(DESCRIPTION) |>
  summarise(TOTAL_AMOUNT = sum(AMOUNT)) |>
  arrange(desc(TOTAL_AMOUNT))
  
```

### Sentence 2

* **Sentence text**: Rep. Matt Gaetz (R-Fla.) was the program’s second-highest overall spender, according to filings released as of Tuesday, with over $30,000 in lodging expenses and more than $11,000 for meals in 2023.
* **Analysis summary**: We found found that Matt Gaetz spent $51,000 on lodging in the year 2023, which is in fact over the listed $30,000. However, the primary lodging code (2107) had just $36,000 expensed to it. We also found over $18,000 in meal expenditures, which does exceed $11,000.
```{r}
matt_gaetz <- rep_expenses |>
  filter(ORGANIZATION == "2023 HON. MATT GAETZ")

matt_gaetz |>
  group_by(DESCRIPTION) |>
  summarize(TOTAL_AMOUNT = sum(AMOUNT)) |>
  arrange(desc(TOTAL_AMOUNT))
```

### Sentence 3

* **Sentence text**: Rep. Jim Banks (R-Ind.), who owns a $1 million home in Virginia, was reimbursed less than $1,500 each month.
* **Analysis summary**: Jim Banks did not, in fact, ever spend more than $1,500 per month in lodging.

```{r}
jim_banks <- rep_expenses |>
  filter(ORGANIZATION == "2023 HON. JIM BANKS")

jim_banks |>
  filter(DESCRIPTION =="LODGING") |>
  mutate(MONTH = month(as.Date(PERFORM.START.DT, format = "%d-%b-%y"))) |>
  group_by(DESCRIPTION) |>
  summarize(AMOUNT, MONTH) |>
  arrange(desc(MONTH))
```

### Sentence 4

* **Sentence text**: Mace, who co-owns a $1,649,000 Capitol Hill townhouse she purchased in 2021 with her then-fiancé, Patrick Bryant, expensed a total of $27,817 in 2023, an average of more than $2,300 a month, according to the data released as of Tuesday. She expensed over $3,000 for lodging in January, March and May, and over $4,000 for lodging in October.
* **Analysis summary**: The data for Mace shows that she expensed closer to $33,00 just for lodging in 2023, and we did find that there were expenses exceeding $3,000 in January, March and May. However, we could not find a single lodging transaction for the month of October, let alone a bill exceeding $4,000

```{r}
nancy_mace <- rep_expenses |>
  filter(ORGANIZATION == "2023 HON. NANCY MACE")

nancy_mace |>
  filter(DESCRIPTION =="LODGING") |>
  mutate(MONTH = month(as.Date(PERFORM.START.DT, format = "%d-%b-%y"))) |>
  group_by(DESCRIPTION) |>
  summarize(AMOUNT, MONTH) |>
  arrange(desc(MONTH))
```

### Sentence 5

* **Sentence text**: Swalwell, who purchased a $1,215,000 home in the Eckington neighborhood of Washington, was reimbursed more than $20,000 for lodging expenses in 2023, according to the data released as of Tuesday. In May 2023, he received $2,838 in lodging reimbursement.
* **Analysis summary**: We were able to confirm that Eric Swalwell spent over $28,000 in 2023 on lodging, and we were able to find the exact transaction matching the $2,838 expense. It is worth noting that the expense was not even his greatest expenditure.
```{r}
eric_swalwell <- rep_expenses |>
  filter(ORGANIZATION == "2023 HON. ERIC SWALWELL")

eric_swalwell |>
  group_by(BUDGET.OBJECT.CODE) |>
  summarize(TOTAL_AMOUNT = sum(AMOUNT)) |>
  arrange(desc(TOTAL_AMOUNT))
```

-30-