---
title: "lab_02"
author: "derek willis"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

For this lab we'll be working with a slice of Maryland voter registration data: records of people who have registration dates this year through Aug. 7. Let's load it.

```{r}
md_voters_2024 <- read_rds("data/lab_02.rds")
```

## Examine Data

Take a look at the data, getting familiar with what it contains. Most of the columns should be self-explanatory, but there are two date columns: STATE_REGISTRATION_DATE and COUNTY_REGISTRATION_DATE. The former represents the date the person registered as a Maryland voter; the latter is when they registered in their current county. They may be the same, but if a voter moves from one county to another they will differ. This is important because if the two dates are the same, that's a pretty good indication that this is a *new* voter, someone who has no history of registering in Maryland. You'll need to keep that definition in mind.

## Answer questions

**Q1** How many voters appear to be *new* registrants since July 21, when President Joe Biden announced he would leave the race, based on the information above? Create a dataframe to hold the only those voters. Which counties have seen the most? Take a look at county population figures here and compare the two in your answer: https://worldpopulationreview.com/us-counties/maryland.

**A1** There were 11,352 new voters registered on or after July 21. Montgomery County saw the most new registrants with 3,404, which stands to reason seeing as it is the most populous county. In fact, the counties with the highest voter registrations are those with the most people.

```{r}
new_md_voters_2024 <- md_voters_2024 |>
  mutate(
    new_voter = case_when(
      day(STATE_REGISTRATION_DATE) >= 21 & month(STATE_REGISTRATION_DATE) >= 07 ~ "YES"
    )
  )

new_md_voters_2024 |>
  filter(new_voter == "YES") |>
  group_by(COUNTY) |>
  summarize(total_regs = n()) |>
  arrange(desc(total_regs))
```

**Q2** A better question is whether the new voters you found in Q1 are an outlier compared to other weeks this year. Using the original dataframe with 2024 registrants, use `mutate()` and lubridate to produce week numbers from the COUNTY_REGISTRATION_DATE (Lubridate has a function that turns a date into a week number. Use that to answer this question). Save that to the original dataframe. Then calculate how many voters are in each week. Describe the distribution of voters groups in a sentence - are the weeks from July 21-August 7 different from the rest, and if so, how?

**A2** The weeks between July 21 and August 7 saw more county voter registrations than most other weeks, except week 24 which had 7,540 registrations

```{r}
new_md_voters_weekly_2024 <- new_md_voters_2024 |>
  mutate(COUNTY_REGISTRATION_DATE = week(COUNTY_REGISTRATION_DATE))

new_md_voters_weekly_2024 |>
group_by(COUNTY_REGISTRATION_DATE) |>
  summarize(total_regs = n()) |>
  arrange(desc(COUNTY_REGISTRATION_DATE))
```

**Q3** Using lubridate, show which month saw the largest number of registrations. Does the answer help shed more light on Q2?
**A3** Yes, there was a spike in registration just before primaries

```{r}
new_md_voters_monthly_2024 <- new_md_voters_2024 |>
  mutate(COUNTY_REGISTRATION_DATE = floor_date(COUNTY_REGISTRATION_DATE, "month"))

new_md_voters_monthly_2024 |>
  group_by(COUNTY_REGISTRATION_DATE) |>
  summarize(total_regs = n()) |>
  arrange(desc(COUNTY_REGISTRATION_DATE))
```

**Q4** Maryland has, officially, five regions: https://www.visitmaryland.org/article/maryland-regions. Modify the original dataframe again to add a `region` column based on the `COUNTY` value using case_when. Are you going to have to do this for every Maryland county plus Baltimore City? Yes, you are. Which regions and parties account for the most new voters?

**A4** The Capital region and Central region account for the most new voters, although admittedly I could not get this function to work for R

```{r}
new_md_voters_2024 <- new_md_voters_2024 |>
  mutate(
    region = case_when(
      COUNTY = "Garrett" && "Allegany" && "Washington" ~ "WESTERN",
      COUNTY = "Frederick" && "Montgomery" && "Prince George's" ~ "CAPITAL",
      COUNTY = "Anne Arundel" && "Baltimore City" && "Baltimore County" && "Carroll" && "Harford" && "Howard" ~ "CENTRAL",
      COUNTY = "Calvert" && "Charles" && "Saint Mary's" ~ "Southern",
      COUNTY = "Kent" && "Queen Anne's" && "Talbot" && "Caroline" && "Dorchester" && "Wicomico" && "Sommerset" && "Worchester" && "Cecil" ~ "EAST",
    )
  )

new_md_voters_2024 |>
  filter(new_voter == "YES") |>
  group_by(region) |>
  summarize(regs = n()) |>
  arrange(desc(region))
```

**Q5** What's an additional question you could ask of this data?

**A5** You could find out whether the ages of the new voters trended lower than other new voters.
